An Accurate Model for SARS-CoV-2 Pooled RT-PCR Test Errors

### 2021-09-14 paper.tex
because FUCK ScholarOne Manuscriptsgit add model_pic.jpg git add model_pic.jpg !

\begin{figure}[H]
  \centering
  \begin{tikzpicture}
    \begin{scope}[every node/.style={circle,thick,draw}]
      \node[shape=circle,draw=black] (A) at (0,0) {$\theta_1=1$};
      \node[shape=circle,draw=black] (B) at (3,0) {$\theta_2=1$};
      \node[shape=circle,draw=black] (D) at (6,0) {$\theta_3=0$};
      %\node[shape=circle,draw=black] (E) at (6,0) {$\theta_4=1$};

      \node[shape=ellipse,draw=black] (F) at (8,5) {False-positive};

      \node[shape=rectangle,draw=black] (G) at (1.5,5)
           {$\Pr(\data=0|\theta)=(1-\sens)^2\spec$};
    \end{scope}

    \begin{scope}[>={Stealth[black]},
        every node/.style={fill=white},
        every edge/.style={draw=black,very thick}]
      \path [->] (A) edge node {$1-\sens$} (2,4.6);
      %\path [->] (B) edge node {$0$} (G);
      %\path [->] (D) edge node {$0$} (G);
      \path [->] (B) edge node {$1-\sens$} (2.5,4.6);
      \path [->] (F) edge node {$\spec$} (3.7,5);
      %\path [-Bar] (D) edge (4.5, 2);
    \end{scope}
  \end{tikzpicture}
  \caption{Illustration of the revised probabilistic model. A pool
    contains individuals $\{1,2,3\}$ with state $\theta=(1,1,0)$
    (i.e., individuals 1 and 2 are infected). A negative pooled test
    ($\data=0$) occurs when three detection paths fail: A
    false-negative occurs for individuals $1$ and $2$, each with
    probability $(1-\sens)$.  Additionally, no false-positive
    detection occurs, with probability $\spec$. Individual $3$ is not
    infected and does not contribute to the probability of the pooled
    test result.}\label{fig:likelihood}
\end{figure}

@@

Interestingly, an increase in infection prevalence monotonically
decreases the scheme false-negative rate, as can also be easily seen
from equation \eqref{eq:sfn}. For the chosen parameter ranges, the increase in
the single test false-negative rates increases the relative error
$\err{exact}$. These effects can be seen in Figure~\ref{fig1} (left
panel), upon conditioning on pool size. Extending the range for $\Sp$
yields no qualitative differences. We further compare $\err{exact}$ to
$\err{common}$, showing the discrepancy changes as a function of both
prevalence and the single test sensitivity (Figure~\ref{fig1}, right
panel).
\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{heatmap_sfn.jpg}
  \caption{Relative increase in Dorfman pooling false-negative rates
    $\err{exact}$. Left: Colors represent $\err{exact}$, the relative
    percentage increase in the scheme false-negative rates relative to
    the single test false-negative rates equation \eqref{eq:erel}. Right:
    colors represent the difference between $\err{common}$ and
    $\err{exact}$. The infection prevalence $q$, is varied on the
    x-axis, while the test sensitivity is varied on the y-axis. Pool
    size $N$, was chosen according to $q$ as in
    \cite{DorfmanYuvalDor}. The left panel shows that $\err{exact}$ is
    largest for low prevalence values $q$. The difference between
    $\err{common}$ and $\err{exact}$ can be as large as 30\%, as seen
    in the right panel. }\label{fig1}
\end{figure}

@@

%% Todos
%% add refs to dilution and paragraph in start

>>>>

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{model_pic.jpg}
  %% \begin{tikzpicture}
  %%   \begin{scope}[every node/.style={circle,thick,draw}]
  %%     \node[shape=circle,draw=black] (A) at (0,0) {$\theta_1=1$};
  %%     \node[shape=circle,draw=black] (B) at (3,0) {$\theta_2=1$};
  %%     \node[shape=circle,draw=black] (D) at (6,0) {$\theta_3=0$};
  %%     %\node[shape=circle,draw=black] (E) at (6,0) {$\theta_4=1$};

  %%     \node[shape=ellipse,draw=black] (F) at (8,5) {False-positive};

  %%     \node[shape=rectangle,draw=black] (G) at (1.5,5)
  %%          {$\Pr(\data=0|\theta)=(1-\sens)^2\spec$};
  %%   \end{scope}

  %%   \begin{scope}[>={Stealth[black]},
  %%       every node/.style={fill=white},
  %%       every edge/.style={draw=black,very thick}]
  %%     \path [->] (A) edge node {$1-\sens$} (2,4.6);
  %%     %\path [->] (B) edge node {$0$} (G);
  %%     %\path [->] (D) edge node {$0$} (G);
  %%     \path [->] (B) edge node {$1-\sens$} (2.5,4.6);
  %%     \path [->] (F) edge node {$\spec$} (3.7,5);
  %%     %\path [-Bar] (D) edge (4.5, 2);
  %%   \end{scope}
  %% \end{tikzpicture}
  \caption{Illustration of the revised probabilistic model. A pool
    contains individuals $\{1,2,3\}$ with state $\theta=(1,1,0)$
    (i.e., individuals 1 and 2 are infected). A negative pooled test
    ($\data=0$) occurs when three detection paths fail: A
    false-negative occurs for individuals $1$ and $2$, each with
    probability $(1-\sens)$.  Additionally, no false-positive
    detection occurs, with probability $\spec$. Individual $3$ is not
    infected and does not contribute to the probability of the pooled
    test result.}\label{fig:likelihood}
\end{figure}

@@

Interestingly, an increase in infection prevalence monotonically
decreases the scheme false-negative rate, as can also be easily seen
from equation \eqref{eq:sfn}. For the chosen parameter ranges, the
increase in the single test false-negative rates increases the
relative error $\err{exact}$. These effects can be seen in
Figure~\ref{fig1} (left panel), upon conditioning on pool
size. Extending the range for $\Sp$ yields no qualitative
differences. We further compare $\err{exact}$ to $\err{common}$,
showing the discrepancy changes as a function of both prevalence and
the single test sensitivity (Figure~\ref{fig1}, right panel).
\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{heatmap_sfn.jpg}
  \caption{Relative increase in Dorfman pooling false-negative rates
    $\err{exact}$. Left: Colors represent $\err{exact}$, the relative
    percentage increase in the scheme false-negative rates relative to
    the single test false-negative rates equation \eqref{eq:erel}. Right:
    colors represent the difference between $\err{common}$ and
    $\err{exact}$. The infection prevalence $q$, is varied on the
    x-axis, while the test sensitivity is varied on the y-axis. Pool
    size $N$, was chosen according to $q$ as in
    \cite{DorfmanYuvalDor}. The left panel shows that $\err{exact}$ is
    largest for low prevalence values $q$. The difference between
    $\err{common}$ and $\err{exact}$ can be as large as 30\%, as seen
    in the right panel.}\label{fig1}
\end{figure}

@@

%% Todos
%% add refs to dilution and paragraph in start
PCR pooling , or simultaneously testing multiple samples, for the
presence of SARS-CoV-2 is increasing in popularity. It is commonly
assumed that including more than a single infected sample in a pooled
test does not increase the probability of a positive outcome. We
challenge this assumption with experimental data. We then analyze the
probability of misidentifying an infected individual and show that it
increases when the prevalence of infection decreases. However, low
infection prevalence is exactly the condition when pooling is most
efficient. We therefore implore the cautious use of pooling.

Relative increase in Dorfman pooling false-negative rates
$\err{exact}$. Left: Colors represent $\err{exact}$, the relative
percentage increase in the scheme false-negative rates relative to the
single test false-negative rates equation \eqref{eq:erel}. Right:
colors represent the difference between $\err{common}$ and
$\err{exact}$. The infection prevalence $q$, is varied on the x-axis,
while the test sensitivity is varied on the y-axis. Pool size $N$, was
chosen according to $q$ as in \cite{DorfmanYuvalDor}. The left panel
shows that $\err{exact}$ is largest for low prevalence values $q$. The
difference between $\err{common}$ and $\err{exact}$ can be as large as
30\%, as seen in the right panel.

Relative increase in Dorfman pooling false-negative rates
$\err{exact}$. Left: Colors represent $\err{exact}$, the relative
percentage increase in the scheme false-negative rates relative to the
single test false-negative rates equation \eqref{eq:erel}. Right:
colors represent the difference between $\err{common}$ and
$\err{exact}$. The infection prevalence $q$, is varied on the x-axis,
while the test sensitivity is varied on the y-axis. Pool size $N$, was
chosen according to $q$ as in \cite{DorfmanYuvalDor}. The left panel
shows that $\err{exact}$ is largest for low prevalence values $q$. The
difference between $\err{common}$ and $\err{exact}$ can be as large as
30\%, as seen in the right panel.
